Visualisations for blog on SCADR website

### 2021-11-15 visualisations_for_blog.Rmd
added plotly plot for geographical variation of place of death, needs a bit more work

```{r, warning=FALSE, message=FALSE}
library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(scico)  # for continuous grayscale palette
library(knitr)  # for printing tables nicely
```

@@

>>>>

```{r, warning=FALSE, message=FALSE}
library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(scico)  # for continuous grayscale palette
library(knitr)  # for printing tables nicely
library(lemon)  # for adding all axis labels to facet_wrap plots
```

@@

# Plotly version of proportions of deaths by place by year by geography

```{r}
library(plotly)  # for interactive plots!
library(htmlwidgets)  # for saving standalone html file

plotly_proportion_of_deaths_by_place_by_hb <-
  (proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  # geom_point(aes(size = place_of_death)) +
  geom_line(
    aes(
      size = place_of_death, 
      group = place_of_death, 
      text = paste0(  # using dummy text aesthetic, construct pairs of variable name: value, separated with line break <br|
        "HB: ", ref_area,
        "<br>Place of death: ", place_of_death,
        "<br>Year: ", year,
        "<br>Proportion: ", scales::percent(proportion_of_total, accuracy = 0.1)
        )
      )
    ) +
  # geom_line(aes(size = place_of_death)) +
  geom_text(data = ~filter(.x, year==2021), aes(x=2021.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = 2015:2021) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80")
    ) +
  labs(
    x = NULL, y= "Proportion of yearly deaths (%)",
    subtitle = paste0(c(paste0("* ",only_2021_text),"Health boards sorted by 2020 population estimate."),collapse="\n"),
    caption = paste0(c(source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2, title = NULL), 
         size = guide_legend(nrow=2, title = NULL)) +
  NULL) %>%
  ggplotly(p = ., tooltip = c("text")) %>%
  layout(legend = list(y = 1.1, orientation = 'h'))  # y=1.1 puts legend above plot

# TODO: add average home death increase in Scotland

plotly_proportion_of_deaths_by_place_by_hb <- plotly_build(plotly_proportion_of_deaths_by_place_by_hb)
for (i in c(1:14, 29:56)) {
  plotly_proportion_of_deaths_by_place_by_hb$x$data[[i]]$visible <- "legendonly"
}

plotly_proportion_of_deaths_by_place_by_hb %>%
  partial_bundle() %>%
  saveWidget(file = paste0(dir_blog,"/plotly_proportion_of_deaths_by_place_by_hb.html"))

```
